---
title: "p8105_hw2_ab4797"
author: "Ani Bilazarian"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

_ Import and Clean Trash Wheel Data

```{r import trash wheel data}
trash_wheel = read_excel("data/trash_wheel_updated.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
rename ("weight" = weight_tons, "volume" = volume_cubic_yards)
```

_ Import and Clean Precipitation data 2018 and adding the variable 'Year'

```{r read and clean precip 18 data}
precip_18 = read_excel("data/trash_wheel_updated.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na() %>% 
    mutate(year = 2018) %>% 
    mutate(month = month.name[month])
```

_ Import and Clean Precipitation data 2017 and adding the variable 'Year'
```{r}
precip_17 = read_excel("data/trash_wheel_updated.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na() %>% 
    mutate (year = 2017) %>% 
    mutate(month = month.name[month]) 

precip_1718 = bind_rows(precip_17, precip_18)
```

_ Dataset Description

This is a publicly available dataset showing us data regardining trash collected from Baltimore's inner harbor. To understand the amount of precipitation from 2017-2018, we combined two data sets. In total there are `r nrow(precip_1718)` observations in the combined resulting datasets. In the dataset precipitation for 2017 there are `r nrow(precip_17)` and in the dataset for 2018 there are `r nrow(precip_18)`. The final dataset has `r ncol(precip_1718)` number of columns, and the key variables represented are Month, Year, and Total Precipitation in inches. The total precipitation in 2018 was `r sum(pull(precip_18, total))` The median number of sportsballs in a dumpster in 2017 was `r median(pull(trash_wheel %>% filter (year == 2017), sports_balls))` 



`` 







