---
title: "Homework 2"
author: "Ani Bilazarian, uni: ab4797"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

_ Import and Clean Trash Wheel Data

```{r import trash wheel data}
trash_wheel = read_excel("data/trash_wheel_updated.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
rename ("weight" = weight_tons, "volume" = volume_cubic_yards)
```

_ Import and Clean Precipitation data 2018 and adding the variable 'Year'

```{r read and clean precip 18 data}
precip_18 = read_excel("data/trash_wheel_updated.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na() %>% 
    mutate(year = 2018) %>% 
    mutate(month = month.name[month])
```

_ Import and Clean Precipitation data 2017 and adding the variable 'Year'
```{r}
precip_17 = read_excel("data/trash_wheel_updated.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
    janitor::clean_names() %>% 
    drop_na() %>% 
    mutate (year = 2017) %>% 
    mutate(month = month.name[month]) 

precip_1718 = bind_rows(precip_17, precip_18)
```

_ Dataset Description

This is a publicly available dataset showing us data regarding trash collected from Baltimore's inner harbor. To understand the amount of precipitation from 2017-2018, we combined two data sets. In total there are `r nrow(precip_1718)` observations in the combined resulting datasets. In the dataset precipitation for 2017 there are `r nrow(precip_17)` and in the dataset for 2018 there are `r nrow(precip_18)`. The final dataset has `r ncol(precip_1718)` number of columns, and the key variables represented are Month, Year, and Total Precipitation in inches. The total precipitation in 2018 was `r sum(pull(precip_18, total))` The median number of sportsballs in a dumpster in 2017 was `r median(pull(trash_wheel %>% filter (year == 2017), sports_balls))` 


# Problem 2

_ Import and Clean Pols data

```{r read in pols data}
pols_month = read_csv("./data/pols-month.csv", col_types = "cciiiciii") %>% 
  janitor::clean_names() %>% 
  separate("mon", into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.integer(month)]) %>% 
  rename("gop" = prez_gop, "dem" = prez_dem) %>% 
  select(-day)
```
_ Import and Clean SNP data

```{r read in SNP data}
snp = read_csv("./data/snp.csv", col_types = "cd") %>% 
  janitor::clean_names() %>% 
  separate("date", into = c("day", "month", "year")) %>% 
  select("year", "month", "day", "close") %>% 
  mutate(month = month.name[as.integer(month)]) %>% 
  select(-day)
```

_ Import and Clean Unemployment data

```{r read in unemployment data}
unemployment_data = read_csv("./data/unemployment.csv", col_types = "cdddddddddddd") %>%     janitor::clean_names() %>% 
    pivot_longer (jan:dec, names_to = "month", values_to = "count") 
```

_ Merge SNP, Pols, and Unemployment Data 

```{r merging SNP into pols}
snp_pols_merge = left_join(pols_month, snp, by = c("month", "year"))
snp_pols_unemployment_merge = left_join(snp_pols_merge, unemployment_data, by = c("month", "year"))
```

_ Dataset Description

This merged data comes from multiple datasets used to indicate when researchers may “p-hack” as a means of finding significant results (a very not cool thing to do!). The “pols-month” data contains observations regarding the number of politicians who are democratic or republican. The S&P data represents a measure of the stock market as a whole, and Unemployment data refers to the number of people unemployed at any given time. After SNP, Unemployment, and Pols datasets were merged, the dataset resulted in `r nrow(snp_pols_unemployment_merge)` total observations. The range of years in this new dataset is `r range(pull(snp_pols_unemployment_merge, year))`. Lastly, key variable names which we want to include are `r names(snp_pols_unemployment_merge)`. 

# PROBLEM 3 

_ Import and Clean Baby Names data

```{r import and clean baby names}
babynames_data_1 = read_csv("./data/Popular_Baby_Names.csv", col_types = "ccccic") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  mutate_all(toupper) %>% 
  rename("year" = year_of_birth, name = childs_first_name) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC")) %>% 
  mutate(ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")) %>% 
 mutate(ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"))  
```

_ Table showing rank of the name "Olivia" 

This table shows the rank in popularity of the name “Olivia” as a female baby name over time according to ethnicities. 

```{r table for Olivia }
babynames_data_1 %>% 
  group_by(year) %>% 
  filter(name == "OLIVIA") %>% 
  select ("year", "ethnicity", "rank") %>% 
  knitr:: kable(digits = 1)
```

_ Table showing most popular name among male children over time

This table shows the most popular name among male children over time according to various ethnicities. The real question we are not asking is... who is naming their child "Liam"? No offense.

```{r table for male names }
babynames_data_1 %>% 
  group_by(year) %>% 
  filter(gender == "MALE", rank == "1") %>% 
  select ("year", "ethnicity", "rank", "name") %>% 
  knitr:: kable(digits = 1)
```

```{r building a scatterplot}
babynames_data_1 %>% 
  filter(ethnicity == "WHITE NON HISPANIC", gender == "MALE", year == "2016") %>% 
ggplot(aes(x = rank, y = count)) + geom_point()
```






